---
title: "Social Network Analysis - Exercise 2"
author: Pair - Yangjia Huang - Thomas Cuna
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(igraph)
library(ggplot2)
library(gridExtra)	
```

On this first exercise, we will learn the basic functions of the Igraph package, a great tool to use Social Network Analysis metrics in our projects. We will use in this first exercise a dataset, "mathoverflow-ints.txt", containing interactions between MathOverflow users with the next information:

- Answers to questions
- Comments to questions
- Comments to answers

You have to complete the code chunks in this document but also analyze the results, extract insights and describe them in an understandable way.

<span style="color:blue">Please, write all your conclusions in blue</span>

Both exercises may be done individually or in pairs; do not forget to write the name of the members of the team at the beginning of this document and also your team number in the name of this file (exercise1_PXX.Rmd)

## Loading data

In this section, the goal is loading the dataset given for this exercise, building the graph object and analyzing basics metrics.

- Build the graph object with all the interactions. Include the timestamp and the type of the edge as edge attributes.
- Describe the values provided by summary function on the graph object.
- ###Validate the graph has been created correctly comparing with the information after reading the file.
- Simplify the graph in order to remove loops and multiple edges. After simplifying the graph, a weight attribute should have been created, counting the number of interactions between two users.

```{r}
#reading data set
data <- read.table(file = "file:///C:/Users/Thomas Cuna/Documents/IE/MBD/3rd Term Classes/Social Network Analysis/exercise1/mathoverflow-ints.txt", sep = ' ')

#add one more column(weight) with value 1
data[,5] <- 1

#rename columns
colnames(data) <- c('from', 'to', 'timestamp', 'type', 'weight')

#creat first igraph object with all the information
g <- graph_from_data_frame(data,directed = T)

#explore the summary of g
summary(g)
```


<span style="color:blue">Answer:

<span style="color:blue">Vertices: 24818

<span style="color:blue">Edges: 506549

<span style="color:blue">Directed: TRUE (DNW -- Directed Named Weighted Graph)

<span style="color:blue">No graph attributes.(V - Vertex, E - Edge, G - Graph)

<span style="color:blue">Vertex attributes: name(v/c - Vertex/Character)

<span style="color:blue">Edge attributes: timestamp (Edge/Numeric), type (Edge/Character), weight (Edge/Numeric).
</span>
```{r}
#simplify the graph: removing multiples and loops but keeping 'weight' attribute
gs <- simplify(g, 
               remove.multiple = T, 
               remove.loops = T, 
               edge.attr.comb = list(timestamp = 'ignore', type = 'ignore', weight = 'sum'))

# show edge of simplified igraph
head(E(gs), 92) 
```

```{r}
# show weight of simplified igraph
head(E(gs)$weight, 92)
```
<span style="color:blue">
Answer:

<span style="color:blue">
Weight values are no longer equal to 1, but is the sum(or aggregate)
</span>

## Analyzing the graph

In this section, we will use Igraph functions to analyze some basic features:

- Is the graph connected? If it is not, compute the connected components and analyze the statistical distribution of the size of the resulting components. Do this considering all the links but also using only the different types of edges. 
- Analyze the statistical distribution of the "weight" attribute created in the previous section and fit it to a power law distribution. Does it fit well enough? 

```{r}
#Are the graphs Connected?
IS_g <- is.connected(g)
IS_gs <- is.connected(gs)

cat(paste(
          paste0("the original graph is connected:", IS_g),
          paste0("the simplified graph is connected:", IS_gs),
          sep = "\n"))
```
<span style="color:blue">
Answer:

<span style="color:blue">
the original graph is connected:FALSE

<span style="color:blue">
the simplified graph is connected:FALSE
</span>

```{r}
#Subsetting original graph by types of graph
g_a2q <- subgraph.edges(graph = g, 
                        eids = which(E(g)$type == "a2q"),
                        delete.vertices = T)

g_c2q <- subgraph.edges(graph = g, 
                        eids = which(E(g)$type == "c2q"),
                        delete.vertices = T)

g_c2a <- subgraph.edges(graph = g, 
                        eids = which(E(g)$type == "c2a"),
                        delete.vertices = T)

#Simplifying subseted graph
gs_a2q <- simplify(g_a2q, remove.multiple = T, remove.loop = T,
                   edge.attr.comb = list(timestamp = 'ignore', type = 'ignore', weight = 'sum'))

gs_c2q <- simplify(g_c2a, remove.multiple = T, remove.loop = T,
                   edge.attr.comb = list(timestamp = 'ignore', type = 'ignore', weight = 'sum'))

gs_c2a <- simplify(g_c2q, remove.multiple = T, remove.loop = T,
                   edge.attr.comb = list(timestamp = 'ignore', type = 'ignore', weight = 'sum'))
```

```{r}
#plotting the distributions for original graph and its subsets
par(mfrow = c(2,2))

plot(density(components(g)$membership), log = 'xy', main = 'Original')
plot(density(components(g_a2q)$membership), log = 'xy', main = 'a2q')
plot(density(components(g_c2a)$membership), log = 'xy', main = 'c2a')
plot(density(components(g_c2q)$membership), log = 'xy', main = 'c2q')
```
```{r}
#plotting the distributions for simplified graph and its subsets
par(mfrow = c(2,2))

plot(density(components(gs)$membership), log = 'xy', main = 'Simplified Original')
plot(density(components(gs_a2q)$membership), log = 'xy', main = 'Simplified a2q')
plot(density(components(gs_c2a)$membership), log = 'xy', main = 'Simplified c2a')
plot(density(components(gs_c2q)$membership), log = 'xy', main = 'Simplified c2q')
```
```{r}
#plotting the distributions of weight for simplified graph and its subsets
par(mfrow = c(2,2))

plot(density(E(gs)$weight), log = 'xy', main = 'Simplified weight')
plot(density(E(gs_a2q)$weight), log = 'xy', main = 'Simplified a2q weight')
plot(density(E(gs_c2a)$weight), log = 'xy', main = 'Simplified c2a weight')
plot(density(E(gs_c2q)$weight), log = 'xy', main = 'Simplified c2q weight')
```


<span style="color:blue">
Answer: 

<span style="color:blue">
Solely looking at the graphs, they all seem to have a strong pattern of power-law distribution, to further justify whether they follow the power law distribution, we run the Kolmogorov-Smirnov test as below.
</span>


```{r}
# power-law distribution?
fp_gsw<-fit_power_law(E(gs)$weight)$KS.p
fp_gsw_a2q<-fit_power_law(E(gs_a2q)$weight)$KS.p
fp_gsw_c2a<-fit_power_law(E(gs_c2a)$weight)$KS.p
fp_gsw_c2q<-fit_power_law(E(gs_c2q)$weight)$KS.p

cat(paste(
  paste0('Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05'),        
  paste0("Simplified Original weight  : ", fp_gsw > 0.05),
  paste0("Simplified a2q weight       : ", fp_gsw_a2q > 0.05),
  paste0("Simplified c2a weight       : ", fp_gsw_c2a > 0.05),
  paste0("Simplified c2q weight       : ", fp_gsw_c2q > 0.05),
          sep = "\n"))
```

Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05

<span style="color:blue">
Answer:

<span style="color:blue">
Simplified Original weight  : FALSE

<span style="color:blue">
Simplified a2q weight       : TRUE

<span style="color:blue">
Simplified c2a weight       : TRUE

<span style="color:blue">
Simplified c2q weight       : TRUE
</span>

## Node degree

A typical analysis when dealing with social networks is studying the degree distribution of the nodes.
- Visualize the statisical distribution of the node degree.
- Again, is it a power-law distribution? Explain why this is the case.

```{r}
#plotting distribution of degree of original graph and its subsets
par(mfrow = c(2,2))

plot(density(degree(g)), log = 'xy', main = 'Original node degree')
plot(density(degree(g_a2q)), log = 'xy', main = 'a2q node degree')
plot(density(degree(g_c2a)), log = 'xy', main = 'c2a node degree')
plot(density(degree(g_c2q)), log = 'xy', main = 'c2q node degree')
```


<span style="color:blue">
Answer: 

<span style="color:blue">
Solely looking at the graphs, they all seem to have a strong pattern of power-law distribution, to further justify whether they follow the power law distribution, we run the Kolmogorov-Smirnov test as below.
</span>
```{r}
# power-law distribution?
fp_gd<-fit_power_law(degree(g))$KS.p
fp_gd_a2q<-fit_power_law(degree(g_a2q))$KS.p
fp_gd_c2a<-fit_power_law(degree(g_c2a))$KS.p
fp_gd_c2q<-fit_power_law(degree(g_c2q))$KS.p


cat(paste(
  paste0('Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05'),        
  paste0("Original node degree  : ", fp_gd > 0.05),
  paste0("a2q node degree       : ", fp_gd_a2q > 0.05),
  paste0("c2a node degree       : ", fp_gd_c2a > 0.05),
  paste0("c2q node degree       : ", fp_gd_c2q > 0.05),
          sep = "\n"))
```


Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05

<span style="color:blue">
Answer:

<span style="color:blue">
Original node degree  : TRUE

<span style="color:blue">
a2q node degree       : FALSE

<span style="color:blue">
c2a node degree       : FALSE

<span style="color:blue">
c2q node degree       : TRUE
</span>
```{r}
#plotting distribution of degree of simplified graph and its subsets
par(mfrow = c(2,2))

plot(density(degree(gs)), log = 'xy', main = 'Simplified node degree')
plot(density(degree(gs_a2q)), log = 'xy', main = 'Simplified a2q node degree')
plot(density(degree(gs_c2a)), log = 'xy', main = 'Simplified c2a node degree')
plot(density(degree(gs_c2q)), log = 'xy', main = 'Simplified c2q node degree')

```

<span style="color:blue">
Answer: 

<span style="color:blue">
Solely looking at the graphs, they all seem to have a strong pattern of power-law distribution, to further justify whether they follow the power law distribution, we run the Kolmogorov-Smirnov test as below.
</span>

```{r}
# power-law distribution?
fp_gsd<-fit_power_law(degree(gs))$KS.p
fp_gsd_a2q<-fit_power_law(degree(gs_a2q))$KS.p
fp_gsd_c2a<-fit_power_law(degree(gs_c2a))$KS.p
fp_gsd_c2q<-fit_power_law(degree(gs_c2q))$KS.p

cat(paste(
  paste0('Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05'),        
  paste0("Simplified node degree    : ", fp_gsd > 0.05),
  paste0("Simplified a2q node degree: ", fp_gsd_a2q > 0.05),
  paste0("Simplified c2a node degree: ", fp_gsd_c2a > 0.05),
  paste0("Simplified c2q node degree: ", fp_gsd_c2q > 0.05),
          sep = "\n"))

```


Follow a power-law distribution? TRUE(Y)/FALSE(N)    p_value > 0.05

<span style="color:blue">
Answer:

<span style="color:blue">
Simplified node degree    : FALSE

<span style="color:blue">
Simplified a2q node degree: FALSE

<span style="color:blue">
Simplified c2a node degree: TRUE

<span style="color:blue">
Simplified c2q node degree: FALSE
</span>


## Building connected graphs

In general, in interaction networks, many connected components turn out after creating the graph. However, most of the nodes typically belong to the same giant component. In this section, the goal is:
- Build a subgraph with all the nodes belonging to the largest connected component.
- Validate the graph has been correctly created.
- Analyze how node degree has changed from the original graph to this new subgraph. Find out which nodes have now a smaller number of connections and explain why.


```{r}
#Create subseting function, but only considering strong connection
largest_comp <- function(graph) {
cl <- components(graph, mode = 'strong')
V(graph)[which.max(cl$csize) == cl$membership]
}

#subgraphs of largest connected component of original graph and its sub graphs by types
g_lcc <- subgraph(g, v = largest_comp(g))
g_a2q_lcc <- subgraph(g_a2q, v = largest_comp(g_a2q))
g_c2q_lcc <- subgraph(g_c2q, v = largest_comp(g_c2q))
g_c2a_lcc <- subgraph(g_c2a, v = largest_comp(g_c2a))

#subgraphs of largest connected component of simplified graph and its sub graphs by types
gs_lcc <- subgraph(gs, v = largest_comp(gs))

gs_a2q_lcc <- subgraph(gs_a2q, v = largest_comp(gs_a2q))
gs_c2q_lcc <- subgraph(gs_c2q, v = largest_comp(gs_c2q))
gs_c2a_lcc <- subgraph(gs_c2a, v = largest_comp(gs_c2a))
```
###Changes of degree
```{r}
#Creating a function to summarize the degree
degree_summary <- function(graph){
  dgr <- degree(graph)
  MAX = max(dgr)
  MEAN = mean(dgr)
  MIN = min(dgr)
  MEDIAN = median(dgr)
  SD = sd(dgr)
  return(list('max' = MAX, 
              'mean' = MEAN, 
              'min' = MIN, 
              'median' = MEDIAN, 
              'standard deviation' = SD))
}

#Creating a dgr_table that takes the summary of the degree of each graph
dgr_table <- data.frame()

for(i in list(g, g_lcc, g_a2q, g_a2q_lcc, g_c2a, g_c2a_lcc, g_c2q, g_c2q_lcc,
           gs, gs_lcc, gs_a2q, gs_a2q_lcc, gs_c2a, gs_c2a_lcc, gs_c2q, gs_c2q_lcc)){
  rw <- as.data.frame(degree_summary(i))
  dgr_table <- rbind(dgr_table, rw)
}

#Creating a dgr_diff_table that calculate the different between every two rows of dgr_table, for example the diff_1 = g_lcc - g
dgr_diff_table <- data.frame(matrix(ncol = 5))
colnames(dgr_diff_table) <- colnames(dgr_table)

for(i in 1:8){
  dgr_diff_table <- rbind(dgr_diff_table, dgr_table[2*i - 1, ])
  dgr_diff_table <- rbind(dgr_diff_table, dgr_table[2*i, ])
  dgr_diff_table <- rbind(dgr_diff_table, (dgr_table[2*i, ] - dgr_table[2*i - 1, ]))}

dgr_diff_table <- dgr_diff_table[-1,]

rownames(dgr_diff_table) <- c('g', 'g_lcc', 'diff_1',
                          'g_a2q', 'g_a2q_lcc', 'diff_2',
                          'g_c2a', 'g_c2a_lcc', 'diff_3',
                          'g_c2q', 'g_c2q_lcc', 'diff_4',
                          'gs', 'gs_lcc', 'diff_5',
                          'gs_a2q', 'gs_a2q_lcc', 'diff_6',
                          'gs_c2a', 'gs_c2a_lcc', 'diff_7',
                          'gs_c2q', 'gs_c2q_lcc', 'diff_8')
#degree difference table
dgr_diff_table
```
```{r}
p1 <- ggplot() +
  geom_density(aes(log(degree(g)), ), colour = 'blue', fill = 'blue', alpha = 0.1) +
  geom_density(aes(log(degree(g_lcc))), colour = 'red', fill = 'red', alpha = 0.1) +
  xlab('log_degree')+ggtitle('Normal')

p2 <- ggplot() +
  geom_density(aes(log(degree(gs)), ), colour = 'blue', fill = 'blue', alpha = 0.1) +
  geom_density(aes(log(degree(gs_lcc))), colour = 'red', fill = 'red', alpha = 0.1) +
  xlab('log_degree')+ggtitle('Simplified')

grid.arrange(p1, p2, nrow = 1)

```

<span style="color:blue">
Based on the above graphs: When the the largest component of the graph is extracted for both the Original and Simplified, we see from the 2 distribution graphs that the log of degree of largest component graph(red) seems to be more normally distributed as compared to the log of the degree of the whole network(blue) because of the outlier nodes with which have less degrees of relation because they have smaller networks.
</span>

# Visualizing shortest paths

On this previous subgraph, you have to compute:

- The shortest path from the node with the highest degree to the rest of the graph.
- From all these paths, select one path with the maximum length.
- Visualize only this path using the plot function.
- Visualize this path within the giant component subgraph.

To this end, use the layout.fruchterman.reingold function to place the nodes in the visualization.

```{r}
#Vertex with highest degree
v_max_dgr <- which(degree(gs_lcc) == max(degree(gs_lcc)))

#Shortest Paths to vertex with highest degree
sps <- shortest_paths(gs_lcc, from = v_max_dgr, to = V(gs_lcc))

#Vertices of Longest Shortest Path to vertex with highest degree
v_max_len <- sps$vpath[which(lengths(sps$vpath) == max(lengths(sps$vpath)))]

#Subgraph for the Longest Shortest Path
lsp_sbg <- subgraph(gs_lcc, v = v_max_len[[1]])

#Plotting lsp_sg
plot(lsp_sbg)
```


<span style="color:blue">
Answer:

<span style="color:blue">
Name of Node with Highest Degree: 290

<span style="color:blue">
Degree: 50

<span style="color:blue">
Number of Vertices of the Longest Shortest Path from vertex 290: 7

<span style="color:blue">
List of the Vertices : 290   3384  2954  48831 57483 58271 58112
</span>

```{r}

#Visuliazing the longest path(in red) within the largest connected component
E(gs_lcc)$color <- rgb(0,0,1,0.01)
E(gs_lcc)$width <- 0.1
E(gs_lcc, path = as.numeric(v_max_len[[1]]))$color <- rgb(1,0,0,1)
E(gs_lcc, path = as.numeric(v_max_len[[1]]))$width <- 1

plot(gs_lcc,
     
     edge.arrow.size=0.01,
     vertex.size = 1,
     vertex.color = rgb(0,0,0,0),
     vertex.frame.color = rgb(0,0,0,0),
     vertex.label.color= rgb(0,0,0,0))
```


<span style="color:blue">
Answer:

<span style="color:blue">
Network: Largest Connected Component

<span style="color:blue">
Red Highlighted path: Longest Shortest Path from vertex 290
